---
title: "Factor Analysis Report for Body Measurements Dataset"
author: "S/17/314"
date: "2023-05-27"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Introduction

  The Body Measurements Dataset is a comprehensive collection of anthropometric measurements captured from a diverse sample of individuals. Anthropometry refers to the measurement of human body dimensions and proportions, and it plays a crucial role in various fields such as health sciences, ergonomics, clothing design, and nutrition research. 
  
Test the hypothesis that the selected factors are sufficient.
  
The objectives of factor analysis are to reduce data dimensionality, identify latent variables, summarize information, explore relationships among variables,
  

# Methodology

## variable Information 
Gender (Male and Female (M=1 & F= 2) (391 Males & 324 Females)
Age (1 year and above)
HeadCircumference (in inches)
ShoulderWidth (in inches)
ChestWidth (in inches)
Belly (in inches)
Waist (in inches)
Hips (in inches)
ArmLength (in inches)
ShoulderToWaist (in inches)
WaistToKnee (in inches)
LegLength (in inches)
TotalHeight - from head to toe (in inches)

Factor analysis is a statistical approach used to identify the underlying factors or dimensions that explain the relationships between a group of observable data. It seeks to simplify the data by lowering its dimensionality and detecting common patterns among the variables.




```{r}
library(tidyverse)
library(janitor)
library(data.table)
library(factoextra)
library(psych)
library(corrplot)
library(ggplot2)
```

```{r}
Body_measurement <- read_csv(file="Data/Body_measurement.csv")
head(Body_measurement)
```
remove empty rows and columns

```{r}
Body_measurement <- Body_measurement %>% remove_empty(c("cols"))
Body_measurement <- Body_measurement %>% drop_na()
```


```{r}
head(Body_measurement)
```

```{r}
describe(Body_measurement)
```
mean of each variable
```{r}
apply(Body_measurement, 2, mean)
```

variance of each variable

```{r}
apply(Body_measurement, 2, var)
```
Correlation Matrix for the data

```{r}
cor(Body_measurement, method = "spearman")

```

```{r}
scaled_df <- apply(Body_measurement, 2, scale)
head(scaled_df)
```

The test measures sampling adequacy for each variable in the model 

```{r}
KMO(scaled_df)
```

Dimention of the dataset
```{r}
dim(scaled_df)
```

Get the covariance matrix of standardized data
```{r}
df_cov <-cov(scaled_df)
df_cov
```


Find the eigen value and eigen vectors of covariance matrix
```{r}
df_eigen <- eigen(df_cov)
df_eigen
```

proportion of variance 

```{r}
PVE <- df_eigen$values / sum(df_eigen$values)
PVE
```
Scree Plot


```{r}
# Create scree plot
plot(PVE, type = "b", xlab = "Factor", ylab = "Proportion of Variance Explained", main = "Scree Plot")

```


`

Factor Analysis from Principal component Method
```{r}
df_st_fa_pc <- fa(df_cov ,nfactors = 3,rotate = "none",n.obs = 252 ,covar = TRUE,fm = "pa")
df_st_fa_pc
```

Get unrotated loadings from PC method
```{r}
df_st_fa_pc$loadings
```


Get unrotated PC loadings as dataframe
```{r}
unrotated_pc_loadings <- as.data.frame(unclass(df_st_fa_pc$loadings))
unrotated_pc_loadings
```

Get unrotated PC method communalities as dataframe
```{r}
unrotated_pc_com <- as.data.frame(unclass(df_st_fa_pc$communality))
unrotated_pc_com
```

Rotate the PC method factor loadings using 'Varimax' method
```{r}
library(GPArotation)
df_st_fa_pc_rotate <- fa(df_cov ,nfactors = 3,rotate = "varimax",n.obs = 252 ,covar = TRUE,fm = 'pa')
df_st_fa_pc_rotate
```
```{r}
df_st_fa_pc_rotate$loadings
```


Roatate factor loadings as a dataframe from PC method
```{r}
rotated_pc_loadings <- as.data.frame(unclass(df_st_fa_pc_rotate$loadings))
rotated_pc_loadings
```
Get rotated communalities in PC method
```{r}
rotated_pc_com <- as.data.frame(unclass(df_st_fa_pc_rotate$communality))
rotated_pc_com
```